Predicting Water Treatment Plant Chemical Doses

### 2025-12-10 Presentation DSCI445.Rmd
Edited data cleaning slide, other.

- Daily averages were taken for raw water data to match the dosing data's daily timestep 
- Merged various datasets into a final cleaned dataset
- Various large outliers (likely caused by equipment malfunction) were removed
- Temperature outliers were interpolated to nearest adjacent values
- Suspended grain size was not used as it was highly variable
- Alkalinity and hardness were not used because of missing data chunks and differing time ranges

@@

```{r, fig.width=7, fig.height=2}

plot_pH <- ggplot(df_lag_incdec, aes(x = Timestamp,
                           y = mean_pH,
                           color = mean_pH)) +
  geom_line() + 
  scale_color_gradient(low = 'blue', high = 'red') +
  theme_minimal() +
  ggtitle("Average pH")

print(plot_pH)

plot_temp <- ggplot(df_lag_incdec, aes(x = Timestamp,
                           y = mean_Temp,
                           color = mean_Temp)) +
  geom_line() + 
  scale_color_gradient(low = 'blue', high = 'red') +
  theme_minimal() +
  ggtitle("Water Temperature")

print(plot_temp)

```{r, fig.width=7, fig.height=2}

plot_cond <- ggplot(df_lag_incdec, aes(x = Timestamp,
                           y = mean_Conductivity,
                           color = mean_Conductivity)) +
  geom_line() + 
  scale_color_gradient(low = 'purple', high = 'yellow') +
  theme_minimal() +
  ggtitle("Water Conductivity")

print(plot_cond)

plot_turb <- ggplot(df_lag_incdec, aes(x = Timestamp,
                           y = mean_Turbidity,
                           color = mean_Turbidity)) +
  geom_line() + 
  scale_color_gradient(low = 'black', high = 'orange') +
  theme_minimal() +
  ggtitle("Water Turbidity")

print(plot_turb)

>>>>

- Dropped Data  
   - Suspended grain size was not used as it was highly variable
   - Alkalinity and hardness were not used because of missing data chunks and differing time ranges

- Processing
   - Daily averages were taken for raw water data to match the dosing data's timestep
   - Significant outliers (likely caused by equipment malfunction) were removed or interpolated
   - Dosing and raw water data merged

@@

```{r, fig.width=7, fig.height=4}

ggplot(df_lag_incdec_long, aes(x = Timestamp, y = Value)) +
  geom_point(size = 0.5, color="blue") +
  facet_wrap(~Characteristics, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Date",
    y = "Value",
    title = "Raw Water Characteristics"
  )
